---
title: "Test_synth_package"
author: "Léonie"
date: "2024-10-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Synth)

```

```{r}
data("basque")
```


Economic effects of conflicts: 
```{r}
dataprep.out <- dataprep(foo = basque, predictors = c("school.illit", "school.prim", "school.med", "school.high", "school.post.high", "invest"), predictors.op = "mean", time.predictors.prior = 1964:1969, special.predictors = list( list("gdpcap", 1960:1969 , "mean"), list("sec.agriculture", seq(1961, 1969, 2), "mean")#available data for predictors before treatments (every year from 1960 to 1969 or every two years from 1961 to 1969)
                                                                                                                                                                                                                                       , list("sec.energy", seq(1961, 1969, 2), "mean"), list("sec.industry", seq(1961, 1969, 2), "mean"),  list("sec.construction", seq(1961, 1969, 2), "mean"), list("sec.services.venta", seq(1961, 1969, 2), "mean"), list("sec.services.nonventa", seq(1961, 1969, 2), "mean"), list("popdens", 1969, "mean")), dependent = "gdpcap", unit.variable = "regionno", unit.names.variable = "regionname", time.variable = "year", treatment.identifier = 17,  controls.identifier = c(2:16, 18),  time.optimize.ssr = 1960:1969, time.plot = 1955:1997)
```


dataprep.out <- dataprep(foo = data, predictors = c("environment", "roads (binary)"...), predictors.op = "mean", time.predictors.prior = to define, dependent = "productivity", unit.variable = "RHC_number", unit.names.variable = "RHC name", time.variable = "year", treatment.identifier = number of the treatment,  controls.identifier = no of the control groups,  time.optimize.ssr = 1960:1969, time.plot = 1955:1997)




ChatGPT: 
- Purpose: The argument is used to define predictors that are not time-invariant, meaning they vary over time. For example, economic indicators like GDP or unemployment rates, which can change yearly, are commonly included as special predictors.

- Structure: The argument takes a list, where each element is a list of three components:

Name of the variable: A character string naming the variable you want to include as a special predictor.
Aggregation function: A function (like mean, sum, median, etc.) that will be used to aggregate this predictor across specified time periods.
Time periods: A vector of the specific time periods for which you want to aggregate this variable.

Average of treated predictor
```{r}
dataprep.out$X1
```


Treated outcome pre-treatment period
```{r}
dataprep.out$Z1
```


Average control predictor
```{r}
dataprep.out$X0
```

control´s outcome pre-treatment period
```{r}
dataprep.out$Z0
```

```{r}
synth.out <- synth(data.prep.obj = dataprep.out, method = "BFGS")#change BFGS by the most adapted algorithm, to get the normality
```
Wikipedia:"The Nelder–Mead method (also downhill simplex method, amoeba method, or polytope method) is a numerical method used to find the minimum or maximum of an objective function in a multidimensional space. It is a direct search method (based on function comparison) and is often applied to nonlinear optimization problems for which derivatives may not be known."

"The Nelder Mead tends to be more successful in estimating the means for all locations than BFGS algorithm, The Nelder Mead tends to be successful rather than BFGS
algorithm in estimating the means of dependent variables
count data with overdispersion. BFGS algorithm tends to
have false convergence in estimating the means of them.
GWMNB is complex and very time consuming, especially
for the simulation study. Nelder mead is more robust related
to the initial value, but it is slower than the BFGS algorithm.
Therefore, this research encourages to find out the fast and
robust algorithms for future studies of geographically
weighted multivariate. The Bayesian or modification of
BFGS can be alternative methods for the next studies" (Yuliani et al.)

"Numerical method like quasi newton BFGS is
commonly used to maximize the Box-Cox function because this method has fast convergence property but it may fail
to convergence in some circumstance. In other hand, direct optimization algorithm like Nelder Mead may still be
convergence in case which quasi newton BFGS fail to convergence but this algorithm generally has slower
convergence property. "" Generally quasi newton BFGS perform more efficient than
Nelder Mead algorithm in terms of number of functions evaluations and number of iteration. However,
the time differences among those two algorithm is slightly different. Simulation using SUSENAS’s data
result in condition where Nelder Mead algorithm is more efficient than quasi-newton BFGS in case of
multivariate data transformation. From the simulation, we can conclude that generally quasi newton
BFGS perform more efficient than Nelder Mead when applied in Box-Cox transformation formula.
However the time elapsed and evaluation function difference between the two algorithm sometime are so
small and become negligible. Therefore, both algorithm can be used side by side to overcome the
weakness of each method, then case which failed to convergent can be minimalized." (Saputra, Angorowatti et al.)

BFGS: For smooth non-linear function

In particular, in the synth() function we solve for a W ∗ that minimizes  ‖X1 − X0W ‖V = √(X1 − X0W )′V (X1 − X0W ) (1)  where V is defined as some (k × k) symmetric and positive semidefinite matrix. The V matrix is introduced to allow different weights to the variables in X0 and X1 depending on their predictive power on the outcome. An optimal choice of V assigns weights that minimize the mean square error of the synthetic control estimator, that is the expectation of (Y1 − Y0W ∗)′(Y1 − Y0W ∗).

Then V ∗ is chosen to minimize  arg min  V ∈V(Z1 − Z0W ∗(V ))′(Z1 − Z0W ∗(V ))where V is the set of all positive definite and diagonal matrices and the weights for the synthetic control are given by W ∗. synth() solves a nested optimization problem that minimizes equation (2), for W ∗(V ) given by equation (1).



```{r}
synth.tables <- synth.tab(dataprep.res = dataprep.out,synth.res = synth.out)
```


Table with the comparison treated and control
```{r}
synth.tables$tab.pred
```
Weight added to the controls
```{r}
synth.tables$tab.w
```
V weight added to the controls
```{r}
synth.tables$tab.v
```
Graph diff treatment control
```{r}
path.plot(synth.res = synth.out, dataprep.res = dataprep.out,Ylab = "real per-capita GDP (1986 USD, thousand)", Xlab = "year",Ylim = c(0, 12), Legend = c("Basque country","synthetic Basque country"), Legend.position = "bottomright")
```

var of the treatment
```{r}
gaps.plot(synth.res = synth.out, dataprep.res = dataprep.out,Ylab = "gap in real per-capita GDP (1986 USD, thousand)", Xlab = "year",Ylim = c(-1.5, 1.5), Main = NA)
```

You can run placebo test using one control as treatment ( remove the real treatment)
